---
title: "Eau Claire COVID"
#author:  "by ____"
date:  "as of `r format(Sys.time()-0*86400, '%B %d, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code:  https://github.com/jjkraker/EauClaireCOVIDsummary
    theme: spacelab
runtime: shiny
---

```{r global, include=FALSE}
library(flexdashboard)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggformula)
library(kableExtra)
library(tidyverse)
library(lubridate)
library(scales)
library(zoo)
library(formattable)
library(gridExtra)
library(ggpubr)
library(mgcv)

source("PlottingFunctions.R")

# ALLdata = read_csv("StateCountydata.csv")
# ORIGdata = read_csv("StateCountydata.csv")
CountyNewdata = read_csv("CountyNewdata.csv")
#CountyNewdata <- CountyNewdata %>% 
#  filter(DATE > "2020/10/08 19:00:00+00")
StateNewdata = read_csv("StateNewdata.csv")
#StateNewdata <- StateNewdata %>% 
#  filter(DATE > "2020/10/08 19:00:00+00")
ALLdata = bind_rows(StateNewdata,CountyNewdata)

ECdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Eau Claire"),DATE))
DAdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Dane"),DATE))
LCdata <- CaseCheck(arrange(filter(ALLdata,NAME=="La Crosse"),DATE))
CFdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Chippewa"),DATE))
DUdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Dunn"),DATE))
BAdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Barron"),DATE))
POdata <- CaseCheck(arrange(filter(ALLdata,NAME=="Portage"),DATE))
WIdata <- CaseCheck(arrange(filter(ALLdata,NAME=="WI"),DATE))

#estimates of populations from https://www.census.gov/quickfacts, unless otherwise noted
ECpop = 104646 
EC20pop = 20930 # about 20% of 104,646 = 20,930 as number in 20s, with students, https://censusreporter.org/profiles/05000US55035-eau-claire-county-wi/
ECprops = c(0.11,0.13,0.20,0.13,0.10,0.12,0.11,0.10)
DApop = 546695
DAprops = c(0.11,0.13,0.19,0.14,0.12,0.12,0.10,0.09)
DA20pop = round(.19*DApop)
DUpop = 45368
DUprops = c(0.11,0.15,0.19,0.11,0.11,0.12,0.11,0.10)
LCpop = 118230
LCprops = c(0.10,0.14,0.18,0.115,0.115,0.12,0.12,0.11)
LC20pop = round(LCprops[3]*LCpop)
CFpop = 64658 
CFprops = c(0.12,0.13,0.11,0.13,0.12,0.15,0.13,0.11)
BApop = 45244 
BAprops = c(0.12,0.12,0.10,0.12,0.11,0.15,0.14,0.14)
WIpop = 5822434 
WIprops = c(0.11,0.13,0.13,0.13,0.12,0.14,0.13,0.11)
WI20pop = 768561 # about 13.2% of 5,822,434 as number in 20s
POpop = 70772

#<h2 style="color:Tomato;">From [WI DHS](https://www.dhs.wisconsin.gov/covid-19/data.htm):  "As of May 27, 2021, the visualizations are using an updated data file that allows corrections due to quality assurance to be counted on the date when a case or death was first reported, rather than affecting the current daily count of cases or deaths." New cases will include both confirmed and probable.</h2>

#UniversityDates = c("2020-08-31","2020-09-14","2020-09-28", "2021-01-31","2021-02-14","2021-02-28")
#UniversityDates = c("2020-05-15","2020-05-22","2020-10-16", "2020-10-23","2020-11-20","2020-11-27")
UniversityDates = c("2020-07-04","2021-07-04")
# fix Date markings
# adjust Date range?
MaskingDates = c("2021-05-13","2021-05-18", "2021-07-27", "2021-07-30")
startDate = "January 01 2021"; startrow = (1:length(WIdata$DATE))[WIdata$DATE == startDate]
endDate = "August 17 2021"; endrow = (1:length(WIdata$DATE))[WIdata$DATE == format(endDate, format="%B %d %Y")]
#endDate = today()-3; endrow = (1:length(WIdata$DATE))[WIdata$DATE == format(endDate, format="%B %d %Y")]

ECdata2021 <- ECdata[startrow:endrow,]
DAdata2021 <- DAdata[startrow:endrow,]
LCdata2021 <- LCdata[startrow:endrow,]
CFdata2021 <- CFdata[startrow:endrow,]
DUdata2021 <- DUdata[startrow:endrow,]
BAdata2021 <- BAdata[startrow:endrow,]
POdata2021 <- POdata[startrow:endrow,]
WIdata2021 <- WIdata[startrow:endrow,]

locations <- c("Eau Claire county", "Chippewa county", "Barron county", 
               "La Crosse county", "Dane county", "state of Wisconsin")
alldataFull <- list(ECdata,CFdata,BAdata,LCdata,DAdata,WIdata)
alldata2021 <- list(ECdata2021,CFdata2021,BAdata2021,LCdata2021,DAdata2021,WIdata2021)

DataUsed <- ECdata2021
LocName <- "Eau Claire county"
```

Daily+Active  {data-orientation=columns}
=====================================  




Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
# add slider for date?

selectInput('len', 'Length', c("full", "2021only"))
selectInput('loc', 'Location', locations)


```

    
Daily  {data-width=400}
-----------------------------------------------------------------------

**Important dates**: 

   * Blue dashed lines:  delineating mask mandate / advice changes from CDC and local health departments:  

   * Purple lines:  denoting July 4 holiday
 

```{r}

isEC <- function() (input$loc == "Eau Claire county")
isCF <- function() (input$loc == "Chippewa county")
isBA <- function() (input$loc == "Barron county")
isLC <- function() (input$loc == "La Crosse county")
isDA <- function() (input$loc == "Dane county")
isWI <- function() (input$loc == "state of Wisconsin")

whichLoc <- function() (input$loc == locations)
typeData <- function() (input$len == "full")

DataUsed <- reactive({
  if (typeData()) {
    if (isEC()) {
      ECdata 
    } else if (isCF()) {
      CFdata
    } else if (isBA()) {
      BAdata
    } else if (isLC()) {
      LCdata
    } else if (isDA()) {
      DAdata
    } else {
      WIdata
    }
    # add other locations
  } else {
    if (isEC()) {
      ECdata2021 
    } else if (isCF()) {
      CFdata2021
    } else if (isBA()) {
      BAdata2021
    } else if (isLC()) {
      LCdata2021
    } else if (isDA()) {
      DAdata2021
    } else {
      WIdata2021
    }
  }
})

# fix with nested ifs
LocName <- renderText({
    if (isEC()) {
      "Eau Claire county" 
    } else if (isCF()) {
      "Chippewa county"
    } else if (isBA()) {
      "Barron county"
    } else if (isLC()) {
      "La Crosse county"
    } else if (isDA()) {
      "Dane county"
    } else {
      "state of Wisconsin"
    }
})

```




```{r,fig.width=3,fig.height=2}

  renderPlot({
    d <- DailyCases(DataUsed(), LocName())
    print(d)
  })
 
```

***

```{r}
  renderPlot({
    d <- DailybyAge(DataUsed(), LocName())
    print(d)
  })
 
```

Spacer {data-width=40}
-------------------------------------


Active Trends {data-width=400}
-------------------------------------

**Estimated Active**, split by Age group:  Estimated recovery time from health experts varies between 10-20 days, with a nontrivial proportion of cases being [long-term / on-going](https://www.theatlantic.com/health/archive/2020/08/long-haulers-covid-19-recognition-support-groups-symptoms/615382/). 

```{r}

  renderPlot({
    d <- ActiveSumbyAge(DataUsed(), LocName(),10)
    print(d)
  })
 
```

***
```{r}

  renderTable({
    d <- PartialTotalTable(DataUsed(), LocName())
    print(d)
  })
 
```
















Trends {data-orientation=columns}
=====================================     


Overall Daily Trends {data-width=400}
-------------------------------------


**Various 7-day average trends**:   Wisconsin tends to have cyclical patterns in testing results, with low spots typically on Mondays (this may be due to lower testing on weekends)


   * Comparing the 7-day averages of Testing and New-Positives can better show rates.
   * Visualizing 7-day averages of Hospitalizations can show lagged effects of diagnosed cases.


```{r}
  renderPlot({
    d <- Daily7day(DataUsed(), LocName())
    print(d)
  })
 
```

***

```{r}
  renderPlot({
    d <- Daily7Hosp(DataUsed(), LocName())
    print(d)
  })
 
```

Spacer {data-width=40}
-------------------------------------

