---
title: "Eau Claire COVID summaries"
#author:  "by Jessica Kraker, ____"
date:  "as of `r format(Sys.time()-2*86400, '%B %d, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code:  https://github.com/jjkraker/EauClaireCOVIDsummary
    theme: spacelab
---

```{r setup, include=FALSE}
library(flexdashboard)
WIdata = read.csv("Data/WIdata.csv")
# names(WIdata)  
library(knitr)
library(dplyr)
library(ggplot2)
library(ggformula)
library(kableExtra)
library(tidyverse)
library(lubridate)

WIdata <- WIdata %>% 
  mutate(DATE = format(as_date(DATE), format = "%B %d %Y")) #%>%   mutate()
source("PlottingFunctions.R")
```


Summary statistics {data-orientation=columns}
=====================================     

Column {.tabset}
-------------------------------------
   
### Wisconsin {data-width=300}
    
```{r}
FullTotalTable(WIdata)
```

Daily counts of COVID cases and outcomes

### Eau Claire

    
```{r}
```
Column {data-width=600}
-------------------------------------
    
### Chart 1
    
```{r}
```

Cumulative Charts {data-orientation=columns}
=====================================  
    

Column {data-width=400}
-------------------------------------
    
### Chart 1
    
```{r}
```
   
Column {.tabset}
-------------------------------------
   
### Chart 2

```{r}
```   
 
### Chart 3
    
```{r}
```

Column {.tabset}
-------------------------------------
   
### Chart 2

```{r}
```   
 
### Chart 3
    
```{r}
```

Daily Charts {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------

### Chart 1

```{r}
```

Row {data-height=400}
-------------------------------------
   
### Chart 2

```{r}
```   
    
### Chart 3

```{r}
```

References {data-orientation=rows}
=====================================     


Column {data-height=100}
-------------------------------------
    
### Data Sources {data-width=200}

   * [Wisconsin Department of Health Services](https://data.dhsgis.wi.gov/datasets/covid-19-historical-data-table/data?where=GEO%20%3D%20%27State%27) with [variable definitions](https://www.dhs.wisconsin.gov/publications/p02677.pdf)
   * [Eau Claire County statistics]( https://coronavirus-and-covid-19-information-hub-eccounty.hub.arcgis.com/)

### Reliable Updated Visuals {data-width=300}

   * Johns Hopkins dashboard (world, national, and county levels)
   * National
   * [Wisconsin DHS website](https://www.dhs.wisconsin.gov/covid-19/county.htm), with interactive graphs; subsettable by county
   * [Re-opening Trends](https://www.nytimes.com/interactive/2020/07/09/us/coronavirus-cases-reopening-trends.html):  that is, why didn't the first wave end?

   
   
Column {data-height=400}
-------------------------------------

###  Visual and code references for selected plots {data-width=200}

*I devised plots that are similar to those found, for example, at the following locations:*

   * [Github repository by coolbaby0208](https://github.com/coolbaby0208/MN-COVID19)
   * Covid Act Now https://covidactnow.org/?s=58173
   * A regional breakdown of the United States COVID-19 curve https://www.reddit.com/r/dataisbeautiful/comments/heglc5/oc_a_regional_breakdown_of_the_united_states/
   * Minnesota updates from Reddit  https://www.reddit.com/r/minnesota/comments/h9j6pj/covid19_update_up_222_230_by_mdh_active_2780/ and https://www.reddit.com/r/minnesota/comments/hfovhg/625_update_34123_positives_360_1406_deaths_9/
   * [Wisconsin updates](https://www.reddit.com/r/CoronavirusWI/comments/hftc25/62520_graphs/) (from Redditor [unknown-and-alone](https://www.reddit.com/user/unknown-and-alone)) 
   * <a href="https://www.reddit.com/r/CoronavirusWI/comments/hftc25/62520_graphs/" target="_blank">Wisconsin updates</a>
   * https://www.reddit.com/r/dataisbeautiful/comments/hjwthw/oc_countylevel_covid19_case_growth_animation_in/

### Scientific References {data-width=500}

[**Asymptomatic Spread**](https://www.pnas.org/content/early/2020/07/02/2008373117)

**Masks** Should I wear a mask in enclosed public spaces?  Unequivocally, **YES**.  References for reasoning:

   * Current [accumulation of over 70 references](https://threader.app/thread/1279144399897866248)
   * Physical distancing, face masks, and eye protection to prevent person-to-person transmission of SARS-CoV-2 and COVID-19: a systematic review and meta-analysis  https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(20)31142-9/fulltext
   * Background for modelling: A modelling framework to assess the likely effectiveness of facemasks in combination with ‘lock-down’ in managing the COVID-19 pandemic   https://royalsocietypublishing.org/doi/10.1098/rspa.2020.0376
   
A couple more in the works:

   * Face mask wearing rate predicts country's COVID-19 death rates https://www.medrxiv.org/content/10.1101/2020.06.22.20137745v1
   * Association of country-wide coronavirus mortality with demographics, testing, lockdowns, and public wearing of masks (Update June 15, 2020) https://www.researchgate.net/publication/342198360_Association_of_country-wide_coronavirus_mortality_with_demographics_testing_lockdowns_and_public_wearing_of_masks_Update_June_15_2020
   * open letter:  https://docs.google.com/forms/d/e/1FAIpQLSel_CEtyTqFjyv_XFFyhPwbH-J6gnfX0YWsiEqlBLE4bsOHGg/viewform?fbclid=IwAR2QaUjDigeGZpGDS0mszJVXqiF4OqBCLlW_9hc7pApBv4WjDlIX1ToXlV4
   * https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(20)31183-1/fulltext?fbclid=IwAR0TSoh6w_Pl54K5ZmmNU6p_cIBo2EoP9BmSxY5Pu2K0aV95IPwk5T53Q6s

https://www.npr.org/sections/health-shots/2020/06/21/880832213/yes-wearing-masks-helps-heres-why

https://www.sfgate.com/science/article/Study-100-face-mask-use-could-crush-second-15333170.php?

https://www.sciencedirect.com/science/article/pii/S2468042720300117?via%3Dihub

http://ftp.iza.org/dp13319.pdf?fbclid=IwAR1jlcs2PjDIvg3AjqxMucwVPfexpxRDOt5T3k24MzNYmTCIE0Resziaeto

https://www.nbcnews.com/news/us-news/no-new-covid-19-cases-reported-after-infected-hair-stylists-n1230346

   